---
title: "Publication-ready t-tests in R"
author: "Rémi Thériault"
date: "February 3, 2022"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Publication-ready t-tests in R}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r global_options, include=FALSE}
library(knitr)

knitr::opts_chunk$set(
  fig.width = 7, fig.height = 7,
  warning = FALSE, message = FALSE
)
knitr::opts_knit$set(root.dir = tempdir())

pkgs <- c("effectsize", "methods", "flextable", "broom", "report")
successfully_loaded <- vapply(pkgs, requireNamespace, FUN.VALUE = logical(1L), quietly = TRUE)
can_evaluate <- all(successfully_loaded)

if (can_evaluate) {
  knitr::opts_chunk$set(eval = TRUE)
  vapply(pkgs, require, FUN.VALUE = logical(1L), quietly = TRUE, character.only = TRUE)
} else {
  knitr::opts_chunk$set(eval = FALSE)
}
```

## Getting started

This function makes it really easy to get all all your *t*-test results in one simple, publication-ready table.

Let's first load the demo data. This data set comes with base `R` (meaning you have it too and can directly type this command into your `R` console).

```{r}
head(mtcars)
```

Load the `rempsyc` package:

```{r}
library(rempsyc)
```

> ***Note:*** If you haven't installed this package yet, you will need to install it via the following command: `install.packages("rempsyc")`.

```{r}
nice_t_test(
  data = mtcars,
  response = "mpg",
  group = "am",
  warning = FALSE
)
```

> ***Note:*** This function relies on the base R `t.test` function, which uses the Welch t-test per default (see why here: https://daniellakens.blogspot.com/2015/01/always-use-welchs-t-test-instead-of.html). To use the Student t-test, simply add the following argument: `var.equal = TRUE`.

Now the best thing about this function is that you can put all your dependent variables of interest in the function call and it will output a sweet, pre-formatted table for your convenience.

```{r}
t.test.results <- nice_t_test(
  data = mtcars,
  response = names(mtcars)[1:6],
  group = "am",
  warning = FALSE
)
t.test.results
```

If we want it to look nice

```{r}
my_table <- nice_table(t.test.results)
my_table
```

> ***Note:*** The *d* is Cohen's *d*, and the 95% CI is the confidence interval of the effect size (Cohen's *d*). *p* is the *p*-value, *df* is degrees of freedom, and *t* is the *t*-value.

### Save table to Word

Let's open (or save) it to word for use in a publication (optional).

```{r, eval = FALSE}
# Open in Word
print(my_table, preview ="docx")

# Save in Word
flextable::save_as_docx(my_table, path = "t-tests.docx")
```

## Special cases

The function can be passed some of the regular arguments of the base `t.test()` function. For example:

### Student t-test (instead of Welch)

```{r}
nice_t_test(
  data = mtcars,
  response = "mpg",
  group = "am",
  var.equal = TRUE
) |>
  nice_table()
```

### One-sided (instead of two-sided)

```{r}
nice_t_test(
  data = mtcars,
  response = "mpg",
  group = "am",
  alternative = "less",
  warning = FALSE
) |>
  nice_table()
```

### One-sample (instead of two-sample)

```{r}
nice_t_test(
  data = mtcars,
  response = "mpg",
  mu = 17,
  warning = FALSE
) |>
  nice_table()
```

### Paired *t* test (instead of independent samples)

Note that for paired *t* tests, you need to use `paired = TRUE`, and you also need data in "long" format rather than wide format (like for the `ToothGrowth` data set). In this case, the `group` argument refers to the participant ID for example, so the same group/participant is measured several times, and thus has several rows.

```{r}
nice_t_test(
  data = ToothGrowth,
  response = "len",
  group = "supp",
  paired = TRUE
) |>
  nice_table()
```

### Multiple comparison corrections

It is also possible to correct for multiple comparisons. Note that only a Bonferroni correction is supported at this time (which simply multiplies the *p*-value by the number of tests). Bonferroni will automatically correct for the number of tests.

```{r}
nice_t_test(
  data = mtcars,
  response = names(mtcars)[1:6],
  group = "am",
  correction = "bonferroni",
  warning = FALSE
) |>
  nice_table()
```

## Integrations

There are other ways to do *t*-tests and format the results properly, should you wish—for example through the `broom` and `report` packages. Examples below.

```{r}
model <- t.test(mpg ~ am, data = mtcars)
```

### `broom` table

```{r}
library(broom)
(stats.table <- tidy(model, conf.int = TRUE))
nice_table(stats.table, broom = "t.test")
```

### `report` table

```{r}
library(report)
(stats.table <- as.data.frame(report(model)))
nice_table(stats.table, report = "t.test")
```

The `report` package provides quite comprehensive tables, so one may request an abbreviated table with the `short` argument.

```{r}
nice_table(stats.table, report = "t.test", short = TRUE)
```

And there you go!

### Thanks for checking in
    
Make sure to check out this page again if you use the code after a time or if you encounter errors, as I periodically update or improve the code. Feel free to contact me for comments, questions, or requests to improve this function at https://github.com/rempsyc/rempsyc/issues. See all tutorials here: https://remi-theriault.com/tutorials.
